The Talking Heads Experiment

### 2015-04-16 chap11/chap11.tex
hyphenation, orphans, widows and some other layout updated

Finally, the robots must build up a lexicon associating names with individuals. When no name
exists yet, the robots can baptise the object with a newly invented name that
will spread in the population through consecutive games because
hearers can acquire the meaning of an unknown name based on feedback
from the speaker after a game. But since a language game is always a
local interaction between only two agents, it is possible that another
agent has invented a different name for the same object which also has
propagated to some extent, and so synonymy (different names for the
same object) is unavoidable. Moreover because of the inherent noise
and unreliability of visual perception and recognition of pointing
gestures as well as different prototype boundaries for different
agents, it is possible that one agent makes an incorrect guess about
the meaning of an unknown name or misinterprets feedback and thus
acquires a different meaning for a name. So homonymy (different
objects for the same name) is unavoidable as well. This raises the
critical question how a shared optimal vocabulary can be reached at
the level of the population without central control or telepathy,
which will require mechanisms for damping synonymy and homonymy at the
individual and population level.\\

\subsection{Semiotic networks}\todo{This used to be set in a \textbackslash bfshape and not in a subsection}
% {\bfshape Semiotic networks}\\

We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of ``semiotic dynamics''. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This lead to the notion of a {\itshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

@@

Given this notion of a network, the key question for designing an effective 
strategy for playing the language game becomes: What are the operators that are building and
changing the semiotic networks in each agent? We focus first on the
vocabulary, i.e. the links between nodes for individual objects $m_i$ and their 
names $f_i$. We use a variant of the strategy used in the Talking Heads experiment. 
Only the hearer changes the score after a game. 
In the case of a successful game, the score $\sigma_{f_i,m_i}$ of the used association is 
increased and its competitors are decreased according to the following
equations with the alignment rate $\gamma = 0.2$: 
\begin{equation}
 \sigma_{m_i,f_i} \leftarrow \sigma_{m_i,f_i} (1-\gamma) + \gamma
\end{equation}
\begin{equation}
 \sigma_{m_j,f_i} \leftarrow \sigma_{m_j,f_i} (1-\gamma) ~~j \neq i
\end{equation}
\begin{equation}
 \sigma_{m_i,f_k} \leftarrow \sigma_{m_i,f_k} (1-\gamma) ~~k \neq i
\end{equation}
A competitor is another individual $m_j$ stored in the hearer's memory 
for the name $f_i$ or another name $f_k$ for the individual 
$m_i$. When the speaker is using a name which is not linked in the network 
of the hearer to any individual, then the hearer adds a new relation between 
the recognised individual (after pointing by the speaker) and this new name. 
This alternative then competes from 
now on with existing associations through the lateral inhibition dynamics. 
We already know from many other experiments that this strategy, when used collectively in 
consecutive games between randomly chosen members of the population, leads to the  
self-organisation of a shared lexicon due to positive feedback.\footnote{
The semiotic dynamics of lateral inhibition has been studied thoroughly by Bart de Vylder as reported in his Ph.D thesis 
(2007 at the VUB AI Lab) and summarised in this paper: \cite{deVylder:2006}.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap11/figs/png-results.pdf}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added.

We have operationalised just one example of such a heuristic. When an object is moving or being moved,
its appearance may change but the observer tracking the object 
still knows that he is dealing with the same object and so he can
exploit this information to fine-tune his semiotic networks. Because the vision system we used on the \textsc{qrio} robots, developed 
by Michael Spranger,\footnote{
This system is described in-depth in: \cite{Spranger:2012vision}.} is able to track an object, it can align object
regions across different images, enriched with top-down predictions of
how an object region will change over time based on Recursive Bayesian
estimation using Kalman filters. The object being moved by 
the experimenter in \figref{fig:png-perception} 
yields two quite different sensory experiences when standing up or lying down which match with 
two different prototypes, but thanks to the tracking heuristic, the semiotic network can 
be rearranged to reflect that they are two views of the same object (see \figref{fig:png-optimisation}).

@@

As a second example of how thinking in terms of semiotic networks is useful,
we turn to an experiment building further on the work on action-word learning \is{action game}
using the \textsc{aibo} reported in the previous chapter.
It was first designed and carried out with the \textsc{qrio} robot in 2007/2008 by Luc Steels and
Michael Spranger with considerable help from Martin Loetzsch\footnote{Papers describing this experiment more fully are: 
\cite{Steels:2008spatial} and \cite{Steels:2008b}.} and 
then ported in 2010 to the MYON, a robot developed by Manfred Hild and his team at the Humboldt 
University in Berlin (see \figref{fig:myon-action}) (\cite{Steels:2012b}). 
A BBC video clip of the experiment can be seen here: \\
https://www.youtube.com/watch?v=lmoXByLkK14
The porting exercise showed that once we discover how to do a certain type of language game on one robot 
we can relatively easily transfer it to another one, after adapting the lowest level sensing and actuation routines.

@@

\begin{figure}
\centerline{\includegraphics[width=0.2\linewidth]{chap11/figs/postures.pdf}\includegraphics[width=0.75\linewidth]{chap11/figs/vision_2.pdf}}
\caption{\label{vision}{\itshape Left:} Example postures of the robot (which is the \textsc{qrio} in this case).
{\itshape Right:} Visual analysis. From left to right we see the source image, the foreground/background distinction and object segmentation (focusing on the upper torso), and the feature signature of this posture represented as a graph connecting the values (on the y-axis)
of the seven centralised normalised moments (on the x-axis).}
\end{figure}

3. The robots need a {\itshape mirror system} that relates perceived actions carried out by others to their own actions. 
The topic of mirror systems has received a lot of attention the past decade because cells were discovered in the brain 
that serve this purpose.\footnote{
An overview of mirror system research and its relevance to the question of the origins of language is given here: 
\cite{Rizzolatti:1998}.}
However the fact that such cells are discovered in the brain 
does not yet tell us how mirror systems are being learned or how they operate. For the action language game, we tried 
several approaches.

In one of them, the robot stands before the mirror (\figref{robot-mirror}) and observes its own 
postures, which generates the data for learning the association between body postures and prototypes. 
The robot selects a posture, and activates the corresponding motor behaviour. This motor behaviour generates a sensory image and 
hence a sensory experience categorised with a particular prototype. Through standard Hebbian learning (which enforces the 
connection between nodes that are simultaneously active) the link between the posture prototype and the motor action gets 
established and progressively enforced.

\begin{figure}
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/robot-mirror.pdf}}
\caption{\label{robot-mirror}The \textsc{qrio} humanoid robot stands before a mirror and performs various motor behaviours thus observing what visual body-images these behaviours generate.}
\end{figure}

@@

The basic idea of a constraint language is that only dataflow is expressed explicitly, not control flow. 
There is a set of slots (variables) and a constraint attempts to reduce the set of possible 
fillers (bindings) for a slot. Constraints are combined in 
a constraint network, based on sharing slots. A solution is found when every slot in the network has unique bindings. 
Usually it requires looping several times through the constraints and often even a rewiring of the constraints to 
find a solution. A numerical example of a set of constraints on natural numbers could be:
\begin{verbatim}
  x = y + z with 2<y<5 and 0<x<5
\end{verbatim}
The constraints expressed here form a network because the same variables x and y (which are the slots of the 
network) appear in different constraints (the sum constraint and two instances of the <-constraint). 
Given this network, constraint propagation can compute easily that x = 4, y = 3, and z = 1.

IRL formalises constraints using prefix notation in the following way. Slots are symbols preceded 
by question marks. Constraints are defined with a list with the first argument the name of the constraint and 
the remaining arguments the slots: 
\begin{verbatim}
 (<constraint> <arg1> ... <arg-n>)
\end{verbatim}
The first argument is called the target of the constraint because it is the most logical 'outcome' 
of applying a constraint, but as shown in the example above, data may flow in all directions. The numerical 
example given earlier translates into the following IRL program. 
\begin{verbatim}
(sum ?x ?y ?z)
(between ?y 2 5)
(between ?x 0 5)
\end{verbatim}

@@

Functional constructions create a unit on top of the word. For example, in the Basic Colour Strategy, 
the colour-category is used in the filter-by-colour-lenient operation (see \figref{fig:basic-strat}). 
The new unit is called colour-unit in \figref{fig:syntactic-graded}, left. 
The meaning specifies the operations to be added to the network, which is in this case: 
\begin{verbatim}
((filter-by-colour-lenient ?s3 ?s2 ?c))
\end{verbatim}
There are two variables that can be linked to other operations with this meaning: ?s3 and ?s2 both of type object-set. 
On the syntactic side, the unit is categorised as a constituent named colour-category.

@@

\figref{fig:syntactic-graded}, right shows another example, namely the structure for an instantiation of the 
graded-membership strategy for the Tarahumara utterance `sita kame'. It is again a Colour-Phrase.
We see that there are now 4 levels of hierarchy. The ColourMembership-Unit pulls together a colour-category and 
a membership-category to form a new colour-category which then becomes a ColourPhrase-Unit again.\\

\subsection{Influence of embodiment}
% {\bfshape Influence of embodiment}\\

@@

Interestingly, the best results in terms of communicative success
are obtained when the agents are allowed to evolve their own colour language, 
partly because there was no limit on the number of prototypes they could introduce. This was better 
than using the set of basic colour prototypes for English, simply because English would turn to more complex expressions
(such as `bluish green') when the basic colour catgories are not effective. Moreover, 
agents still achieve better results compared to the English colour system when they do not need to stick 
dogmatically to it. Indeed this shows why alignment, which is widely observed
in human dialogues,\footnote{See: \cite{Garrod:1987} and \cite{Galantucci:2005}.}
is a good strategy for achieving more successful communication. \\

>>>>

Finally, the robots must build up a lexicon associating names with individuals. When no name
exists yet, the robots can baptise the object with a newly invented name that
will spread in the population through consecutive games because
hearers can acquire the meaning of an unknown name based on feedback
from the speaker after a game. But since a language game is always a
local interaction between only two agents, it is possible that another
agent has invented a different name for the same object which also has
propagated to some extent, and so synonymy (different names for the
same object) is unavoidable. Moreover because of the inherent noise
and unreliability of visual perception and recognition of pointing
gestures as well as different prototype boundaries for different
agents, it is possible that one agent makes an incorrect guess about
the meaning of an unknown name or misinterprets feedback and thus
acquires a different meaning for a name. So homonymy (different
objects for the same name) is unavoidable as well. This raises the
critical question how a shared optimal vocabulary can be reached at
the level of the population without central control or telepathy,
which will require mechanisms for damping synonymy and homonymy at the
individual and population level.

\subsection{Semiotic networks}\todo{This used to be set in a \textbackslash bfshape and not in a subsection}
We have already seen from earlier experiments that the solution to these various issues lies in setting up
the right kind of ``semiotic dynamics''. In the present case, this dynamics should gradually coordinate
sensory experiences, prototypical views, individuals, 
and names, both within a single individual and across the
population. This lead to the notion of a {\itshape semiotic network}, which was a significant conceptual advance 
in the methodology of conceiving and carrying out language game experiments.

@@

Given this notion of a network, the key question for designing an effective 
strategy for playing the language game becomes: What are the operators that are building and
changing the semiotic networks in each agent? We focus first on the
vocabulary, i.e. the links between nodes for individual objects $m_i$ and their 
names $f_i$. We use a variant of the strategy used in the Talking Heads experiment. 
Only the hearer changes the score after a game. 
In the case of a successful game, the score $\sigma_{f_i,m_i}$ of the used association is 
increased and its competitors are decreased according to the following
equations with the alignment rate $\gamma = 0.2$: 
\begin{equation}
 \sigma_{m_i,f_i} ≤ftarrow \sigma_{m_i,f_i} (1-\gamma) + \gamma
\end{equation}
\begin{equation}
 \sigma_{m_j,f_i} ≤ftarrow \sigma_{m_j,f_i} (1-\gamma) ~~j \neq i
\end{equation}
\begin{equation}
 \sigma_{m_i,f_k} ≤ftarrow \sigma_{m_i,f_k} (1-\gamma) ~~k \neq i
\end{equation}
A competitor is another individual $m_j$ stored in the hearer's memory 
for the name $f_i$ or another name $f_k$ for the individual 
$m_i$. When the speaker is using a name which is not linked in the network 
of the hearer to any individual, then the hearer adds a new relation between 
the recognised individual (after pointing by the speaker) and this new name. 
This alternative then competes from 
now on with existing associations through the lateral inhibition dynamics. 
We already know from many other experiments that this strategy, when used collectively in 
consecutive games between randomly chosen members of the population, leads to the  
self-organisation of a shared lexicon due to positive feedback.\footnote{
The semiotic dynamics of lateral inhibition has been studied thoroughly by Bart de Vylder as reported in his Ph.D thesis 
(2007 at the VUB AI Lab) and summarised in this paper: \cite{deVylder:2006}.}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.95\textwidth]{chap11/figs/png-results.pdf}}
\caption{\label{fig:png-results} 
{\itshape Left:} Results for Proper Naming games in a population of 10 agents that have to name 15 individual objects. 
We see that a vocabulary establishes
itself after agents have played a few hundred games. There is again the typical peak until alignment kicks in. 
The number of individuals is larger than 15, implying that agents name prototypical views rather than 
individuals. {\itshape Right:} Results after the addition  
of a heuristic based on tracking an object over time. We see that there are fewer individuals 
than stored prototypical views, which means that agents have improved their ability to recognise object 
identity. }
\end{figure}
With these mechanisms, the population reaches a high level of communicative success (above 90 \%). 
However the average number of individuals in the agents' semiotic networks
is much larger than the number of distinctive objects
introduced in the experiment (see \figref{fig:png-results}, left). Apparently, agents are naming prototypical views of
individual objects instead of the individuals themselves. So we did not adequately address 
how a robot learns to interrelate different views. To do this, various heuristics need to be added. 
\clearpage
We have operationalised just one example of such a heuristic. When an object is moving or being moved,
its appearance may change but the observer tracking the object 
still knows that he is dealing with the same object and so he can
exploit this information to fine-tune his semiotic networks. Because the vision system we used on the \textsc{qrio} robots, developed 
by Michael Spranger,\footnote{
This system is described in-depth in: \cite{Spranger:2012vision}.} is able to track an object, it can align object
regions across different images, enriched with top-down predictions of
how an object region will change over time based on Recursive Bayesian
estimation using Kalman filters. The object being moved by 
the experimenter in \figref{fig:png-perception} 
yields two quite different sensory experiences when standing up or lying down which match with 
two different prototypes, but thanks to the tracking heuristic, the semiotic network can 
be rearranged to reflect that they are two views of the same object (see \figref{fig:png-optimisation}).

@@

As a second example of how thinking in terms of semiotic networks is useful,
we turn to an experiment building further on the work on action-word learning \is{action game}
using the \textsc{aibo} reported in the previous chapter.
It was first designed and carried out with the \textsc{qrio} robot in 2007/2008 by Luc Steels and
Michael Spranger with considerable help from Martin Loetzsch\footnote{Papers describing this experiment more fully are: 
\cite{Steels:2008spatial} and \cite{Steels:2008b}.} and 
then ported in 2010 to the MYON, a robot developed by Manfred Hild and his team at the Humboldt 
University in Berlin (see \figref{fig:myon-action}) (\cite{Steels:2012b}). 
A BBC video clip of the experiment can be seen at: \url{https://youtu.be/lmoXByLkK14}.
The porting exercise showed that once we discover how to do a certain type of language game on one robot 
we can relatively easily transfer it to another one, after adapting the lowest level sensing and actuation routines.

@@

3. The robots need a {\itshape mirror system} that relates perceived actions carried out by others to their own actions. 
The topic of mirror systems has received a lot of attention the past decade because cells were discovered in the brain 
that serve this purpose.\footnote{
An overview of mirror system research and its relevance to the question of the origins of language is given here: 
\cite{Rizzolatti:1998}.}
However the fact that such cells are discovered in the brain 
does not yet tell us how mirror systems are being learned or how they operate. For the action language game, we tried 
several approaches.
\enlargethispage{1\baselineskip}
In one of them, the robot stands before the mirror (\figref{robot-mirror}) and observes its own 
postures, which generates the data for learning the association between body postures and prototypes. 
The robot selects a posture, and activates the corresponding motor behaviour. This motor behaviour generates a sensory image and 
hence a sensory experience categorised with a particular prototype. Through standard Hebbian learning (which enforces the 
connection between nodes that are simultaneously active) the link between the posture prototype and the motor action gets 
established and progressively enforced. 
\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=0.2\linewidth]{chap11/figs/postures.pdf}\includegraphics[width=0.75\linewidth]{chap11/figs/vision_2.pdf}}
\caption{\label{vision}{\itshape Left:} Example postures of the robot (which is the \textsc{qrio} in this case).
{\itshape Right:} Visual analysis. From left to right we see the source image, the foreground/background distinction and object segmentation (focusing on the upper torso), and the feature signature of this posture represented as a graph connecting the values (on the y-axis)
of the seven centralised normalised moments (on the x-axis).}
\end{figure}

\begin{figure}[p]
\centerline{\includegraphics[width=0.6\linewidth]{chap11/figs/robot-mirror.pdf}}
\caption{\label{robot-mirror}The \textsc{qrio} humanoid robot stands before a mirror and performs various motor behaviours thus observing what visual body-images these behaviours generate.}
\end{figure}

@@

The basic idea of a constraint language is that only dataflow is expressed explicitly, not control flow. 
There is a set of slots (variables) and a constraint attempts to reduce the set of possible 
fillers (bindings) for a slot. Constraints are combined in 
a constraint network, based on sharing slots. A solution is found when every slot in the network has unique bindings. 
Usually it requires looping several times through the constraints and often even a rewiring of the constraints to 
find a solution. A numerical example of a set of constraints on natural numbers could be:
\begin{verbatim}
  x = y + z with 2<y<5 and 0<x<5
\end{verbatim}
The constraints expressed here form a network because the same variables x and y (which are the slots of the 
network) appear in different constraints (the sum constraint and two instances of the <-constraint). 
Given this network, constraint propagation can compute easily that x = 4, y = 3, and z = 1. 
\enlargethispage{2\baselineskip}
IRL formalises constraints using prefix notation in the following way. Slots are symbols preceded 
by question marks. Constraints are defined with a list with the first argument the name of the constraint and 
the remaining arguments the slots: 
\begin{verbatim}
 (<constraint> <arg1> ... <arg-n>)
\end{verbatim}
The first argument is called the target of the constraint because it is the most logical 'outcome' 
of applying a constraint, but as shown in the example above, data may flow in all directions. The numerical 
example given earlier translates into the following IRL program. 
\begin{verbatim}
(sum ?x ?y ?z)
(between ?y 2 5)
(between ?x 0 5)
\end{verbatim}

@@

Functional constructions create a unit on top of the word. For example, in the Basic Colour Strategy, 
the colour-category is used in the filter-by-colour-lenient operation (see \figref{fig:basic-strat}). 
The new unit is called colour-unit in \figref{fig:syntactic-graded}, left. 
The meaning specifies the operations to be added to the network, which is in this case: 
\begin{verbatim}
((filter-by-colour-lenient ?s3 ?s2 ?c))
\end{verbatim}
\clearpage
There are two variables that can be linked to other operations with this meaning: ?s3 and ?s2 both of type object-set. 
On the syntactic side, the unit is categorised as a constituent named colour-category.

@@

\figref{fig:syntactic-graded}, right shows another example, namely the structure for an instantiation of the 
graded-membership strategy for the Tarahumara utterance `sita kame'. It is again a Colour-Phrase.
We see that there are now 4 levels of hierarchy. The ColourMembership-Unit pulls together a colour-category and 
a membership-category to form a new colour-category which then becomes a ColourPhrase-Unit again.
\subsection{Influence of embodiment}
% {\bfshape Influence of embodiment}\\

@@

Interestingly, the best results in terms of communicative success
are obtained when the agents are allowed to evolve their own colour language, 
partly because there was no limit on the number of prototypes they could introduce. This was better 
than using the set of basic colour prototypes for English, simply because English would turn to more complex expressions
(such as `bluish green') when the basic colour catgories are not effective. Moreover, 
agents still achieve better results compared to the English colour system when they do not need to stick 
dogmatically to it. Indeed this shows why alignment, which is widely observed
in human dialogues,\footnote{See: \cite{Garrod:1987} and \cite{Galantucci:2005}.}
is a good strategy for achieving more successful communication.
